## Part II: Parameter estimation

### Problem 4

```{r}

set.seed(17)
id <- 17
theta <- id/10
M <- 1000
N <- 1000
poisson_distribution <- matrix(rpois(N*M, lambda=theta), nrow=M)
```

#### Method 2

Lets introduce Z-rv
$$Z:= \frac{\sqrt{n} \cdot (\bar{X} - \theta) }{\theta} \sim N(0 , 1)
$$
to the estimation of $\theta$ using following:

$$
P(|\theta - \bar{X}| \le \frac{z_{\beta} \cdot \theta}{\sqrt{n}}) = P(|Z| \le z_{\beta}) = 2\beta-1
$$
We also know that $2\beta-1 = 1 - \alpha$ so $\beta = 1 - \alpha/2$
Below calculations are helping as to limit $\theta$ from both sides.
If we use the ability of R language we can estimate the theta by using properties of modulus:
$$
\bar{X} - \theta \le \frac{\theta \cdot z_{1 - \frac{\alpha}{2}}} {\sqrt{n}})\\
\theta - \bar{X} \le \frac{\theta \cdot z_{1 - \frac{\alpha}{2}}} {\sqrt{n}})\\
\theta \geq \bar{X} - \frac{\theta \cdot z_{1 - \frac{\alpha}{2}}} {\sqrt{n}})\\
\theta \le \bar{X} + \frac{\theta \cdot z_{1 - \frac{\alpha}{2}}} {\sqrt{n}})\\
$$
And use it in our estimation function. Therefore, we can adjust the equation
$$
P((|\bar{X} - \theta|) \le \frac{\theta \cdot z_{1 - \frac{\alpha}{2}}} {\sqrt{n}}) = P((|\frac{\bar{X}}{\theta}-1|) \le \frac{ z_{1 - \frac{\alpha}{2}}} {\sqrt{n}}) \\= P(1 - \frac{z_{1 - \frac{\alpha}{2}}}{\sqrt{n}}  \le \frac{\bar{X}}{\theta} \le 1 + \frac{z_{1 - \frac{\alpha}{2}}}{\sqrt{n}}) = \\ P (\frac{\sqrt{n} \cdot \bar{X}}{\sqrt{n} + z_{1 - \frac{\alpha}{2}}}  \le \theta \le  \frac{\sqrt{n} \cdot \bar{X}}{\sqrt{n} + z_{1-\frac{\alpha}{2}}}) \\ 1 - \alpha = [\frac{\sqrt{n} \cdot \bar{X}}{\sqrt{n} + z_{1 - \frac{\alpha}{2}}} ; \frac{\sqrt{n} \cdot \bar{X}}{\sqrt{n} + z_{1-\frac{\alpha}{2}}}]
$$
There we get an estimation fot theta using METHOD 2

```{r}

estimation_method2 <- function(data, alpha) {
  n <- length(data)
  return (c(mean(data) - (theta * qnorm(1 - alpha / 2)) / sqrt(n),
            mean(data) + (theta * qnorm(1 - alpha / 2)) / sqrt(n)))
}


cat("METHOD 2", "\n")
for (alpha in c(.01, .05, .1))
{
    func <- function (data_base) estimation_method2(data_base, alpha)
    lst_of_bounds <- apply(poisson_distribution, 1, func)
    cat("For confidence level", 1 - alpha, "\n")
    cat("  the fraction of CI's containing the paraneter is",
        mean(theta >= lst_of_bounds[1,] & theta <= lst_of_bounds[2,]), "\n", sep = " ")
}
```
#### Method 3

Solve the double inequality that was given in a task to calculate confidence level with unknown variance
$$
P(|\bar{X} - \theta| \le \frac{z_{1-\frac{\alpha}{2}} \cdot \theta}{\sqrt{n}})
$$
$$
P(|\frac{\bar{X}}{\theta} - 1| \le \frac{z_{1-\frac{\alpha}{2}}}{\sqrt{n}})
$$

$$
P(1-\frac{z_{1-\frac{\alpha}{2}}}{\sqrt{n}} \le \frac{\bar{X}}{\theta} \le \frac{z_{1-\frac{\alpha}{2}}}{\sqrt{n}}+1)
$$
Lets get the same denominator

$$
P(\frac{\sqrt{n}-z_{1-\frac{\alpha}{2}}}{\sqrt{n}} \le \frac{\theta}{\bar{X}} \le \frac{\sqrt{n}+z_{1-\frac{\alpha}{2}}}{\sqrt{n}})
$$
Multiply the inequality by sample mean and flip the numbers to obtain the theta bounds:


$$
P(\bar{X} \cdot  \frac{\sqrt{n}}{\sqrt{n}+z_{1-\frac{\alpha}{2}}} \le \theta \le \bar{X} \cdot  \frac{\sqrt{n}}{\sqrt{n}-z_{1-\frac{\alpha}{2}}})
$$

Confidence interval of level $1-\alpha$ :

$$
[\bar{X} \cdot  \frac{\sqrt{n}}{\sqrt{n}+z_{1-\frac{\alpha}{2}}};  \bar{X} \cdot  \frac{\sqrt{n}}{\sqrt{n}-z_{1-\frac{\alpha}{2}}}]
$$
```{r}
cat("METHOD 3", "\n")
estimation_method3 <- function(data, alpha) {
  n <- length(data)
  return (c((mean(data) * sqrt(n)) / (qnorm(1 - alpha / 2) + sqrt(n)),
            (mean(data) * sqrt(n)) / (sqrt(n) - qnorm(1-alpha/2))))
}

#Method 3
for (alpha in c(.01, .05, .1))
{
    func <- function (data_base) estimation_method3(data_base, alpha)
    lst_of_bounds <- apply(poisson_distribution, 1, func)
    cat("For confidence level", 1 - alpha, "\n")
    cat("  the fraction of CI's containing the paraneter is",
        mean(theta >= lst_of_bounds[1,] & theta <= lst_of_bounds[2,]), "\n", sep = " ")
}
```
#### Method 4

Using lecture notes we introduce $T=(\bar X-\mu)\frac{\sqrt n}{S}\sim\mathcal{T}_{n-1}$

And using basic calculations described in Ross textbook we get the inequality
$$
\bar X-t^{(n-1)}_{1-\alpha/2}\frac{S}{\sqrt n} \le \theta \le \bar X+t^{(n-1)}_{1-\alpha/2}\frac{S}{\sqrt n}
$$
Which would be such interval confidence interval for $100(1-\alpha)$-:

$$
[\bar X-t^{(n-1)}_{1-\alpha/2}\frac{S}{\sqrt n},\bar X+t^{(n-1)}_{1-\alpha/2}\frac{S}{\sqrt n}]
$$
```{r}

estimation_method4 <- function(data, alpha) {
  n <- length(data)
  return (c(mean(data) - qt(1 - alpha / 2, n - 1) * sd(data) / sqrt(n),
            mean(data) + qt(1 - alpha / 2, n - 1) * sd(data) / sqrt(n)))
}
cat("METHOD 4", "\n")
for (alpha in c(.01, .05, .1))
{
    func <- function (data_base) estimation_method4(data_base, alpha)
    lst_of_bounds <- apply(poisson_distribution, 1, func)
    cat("For confidence level", 1 - alpha, "\n")
    cat("  the fraction of CI's containing the paraneter is",
        mean(theta >= lst_of_bounds[1,] & theta <= lst_of_bounds[2,]), "\n", sep = " ")
}
```



